VFTS682: a confirmed dynamical ejection?

### 2018-05-09 VFTS682.tex
section on generic surroundings done, fixed errors

How do stars form is one longstanding question in astrophysics
\cite{lada:03, zinnecker:07}. It is particularly difficult for massive stars, because these are intrinsically rare
\citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, evolve fast, and
remain enshrouded in their parent cloud during the formation
process. Moreover, observations of young massive stars reveal a
complicated multipliticy structure which requires
explanation \citep[][]{kobulnicky:07,mason:09,sana:11,sana:12,kiminki:12,chini:12,kobulnicky:14,almeida:17,demarco:17}. Understanding massive star formation, possibly as a
function of metallicity, is a key question given the present and upcoming
transient survey \citep[e.g., LSST, BlackGem, LIGO/Virgo O3][]{} which
will reveal transients associated to massive stars
evolution and death.

@@

In this study, we combine the radial velocity measurements from the
VFTS survey \citep[][]{evans:11} with the proper motion from Gaia DR2
to reconstruct the three-dimensional velocity of VFTS682, and test the
null hypothesis that this star was ejected from R136. \todo{check the
  following} Our results indicate that R136 is the likely origin of
this star, and therefore isolated star formation is \emph{not}
required to explain it. However, we find a mild discrepancy between the
apparent age of VFTS682, its kinematic age, and the age estimates for
R136.

In \Secref{sec:sample}, we describe the data selection and validation
process for VFTS 682 and for stars in the surrounding used to
define the local reference frame. \Secref{sec:results} presents our main findings. We conclude
by discussing the implications for theories of
star formation, N-body interactions, and binary evolution in
\Secref{sec:discussion}.

@@

\subsection{Is it a runaway star?}

We first address the question of whether VFTS682 is a typical star
from the kinematic point of view, or whether it is a runaway star with
a significantly large peculiar velocity compared to its surrounding population. The former is what should
be expected if it formed where we observe it today, in relative
isolation from other massive stars.

Using the 437 stars selected as the ``surrounding population'' in
\Secref{sec:sample}, we find averaged proper motion components of
$\mu_\mathrm{RA} = 1.695\pm0.003\,\mathrm{mas\ yr^{-1}}$ and
$\mu_\mathrm{DE} = 0.691\pm0.003\,\mathrm{mas\ yr^{-1}}$. Subtracting these values from the
proper motions of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
components of proper motion of the star relative to the surrounding region
$\mu^\mathrm{sur}_\mathrm{RA} = 0.14\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu^\mathrm{sur}_\mathrm{DE} =
0.09\pm 0.08\,\mathrm{mas\ yr^{-1}}$. These can be converted in the
the components of the transverse velocity $v^\mathrm{sur}_\mathrm{RA}=35\pm17\,\kms$,
$v^\mathrm{sur}_\mathrm{DE}=22\pm19\,\kms$, assuming a distance of
50\,kpc (we do not account for the uncertainty in the distance
estimate when propagating errors). The radial velocity from
\cite{bestenlehner:11} then gives the third component along
the line of sight, allowing us to calculate the speed of the star:

\begin{equation}
  \label{eq:speed}
  v = \sqrt{\left(v^\mathrm{sur}_\mathrm{RA}\right)^2
    +\left(v^\mathrm{sur}_\mathrm{DE}\right)^2+\left(\delta v_\mathrm{rad}\right)^2} = 61 \pm 1
  \, \kms \ .
\end{equation}
This value for the peculiar 3D velocity of VFTS682 make it a ``bona
fide'' runaway star.

\todo{errors are wrong}

\subsection{Does it come from the R136 cluster?}
The inset plot in \Figref{fig:main} shows a zoom in of the region
around R136 which includes VFTS682. To test the hypothesis that the
star was indeed ejected from this region, we check the relative
orientation of its proper motion. The error cone, obtained by
propagating the errors on the proper motion components indicates that
R136 is indeed the most likely origin, as expected by \cite{fujii:11,
  banerjee:12}.

>>>>

How do stars form is one longstanding question in astrophysics
\cite{lada:03, zinnecker:07}. It is particularly difficult for massive stars, because these are intrinsically rare
\citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, evolve fast, and
remain enshrouded in their parent cloud during the formation
process. Moreover, observations of young massive stars reveal a
complicated multipliticy structure which requires
explanation \citep[][]{kobulnicky:07,mason:09,sana:11,sana:12,kiminki:12,chini:12,kobulnicky:14,almeida:17,demarco:17}. Understanding massive star formation, possibly as a
function of metallicity, is a key question given the present and upcoming
transient survey \citep[e.g., LSST, BlackGem, LIGO/Virgo O3][]{}\todo{ref} which
will reveal transients associated to massive stars
evolution and death.

@@

In this study, we combine the radial velocity measurements from the
VFTS survey \citep[][]{evans:11} with the proper motion from Gaia DR2
to reconstruct the three-dimensional velocity of VFTS682, and test the
hypothesis that this star was ejected from R136. We discuss in
\Secref{sec:sample} the data for VFTS682, and the selection of stars
used to define a local reference frame. Our results
indicate that R136 is a bona fide runaway star (\Secref{sec:runaway}), therefore isolated star formation is \emph{not}
required to explain it. We also find that a dynamical ejection from
R136 is compatible with the direction of its velocity
vector. \todo{double check the following}
However, we find a mild discrepancy between the apparent age of
VFTS682, its kinematic age, and the age estimates for R136 (\Secref{sec:r136_origin}).  We conclude
with a very brief discussion on the implications for theories of
star formation, N-body interactions, and binary evolution in
\Secref{sec:discussion}.

@@

\subsection{Is it a runaway star?}
\label{sec:runaway}
We first address the question of whether VFTS682 is a typical star
from the kinematic point of view, or whether it is a runaway star with
a significantly large peculiar velocity compared to its surrounding population. The former is what should
be expected if it formed where we observe it today, in relative
isolation from other massive stars.

Using the 437 stars selected as the ``surrounding population'' in
\Secref{sec:sample} and shown in blue in \Figref{fig:main}, we find averaged proper motion components of
$\langle\mu_\mathrm{RA}\rangle = 1.695\pm0.003\,\mathrm{mas\ yr^{-1}}$ and
$\langle\mu_\mathrm{DE}\rangle = 0.691\pm0.003\,\mathrm{mas\ yr^{-1}}$. Subtracting these values from the
proper motions of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
components of proper motion of the star relative to the surrounding region
$\mu^\mathrm{sur}_\mathrm{RA} = 0.15\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu^\mathrm{sur}_\mathrm{DE} =
0.09\pm 0.08\,\mathrm{mas\ yr^{-1}}$. %% note that the error budget is
% dominated by VFTS682
These can be converted in the
the components of the transverse velocity $v^\mathrm{sur}_\mathrm{RA}=35\pm17\,\kms$,
$v^\mathrm{sur}_\mathrm{DE}=22\pm19\,\kms$, assuming a distance of
50\,kpc (we do not account for the uncertainty in the distance
estimate when propagating errors). The radial velocity from
\cite{bestenlehner:11} then gives the third component along
the line of sight, allowing us to calculate the speed of the star:

\begin{equation}
  \label{eq:speed}
  v = \sqrt{\left(v^\mathrm{sur}_\mathrm{RA}\right)^2
    +\left(v^\mathrm{sur}_\mathrm{DE}\right)^2+\left(\delta v_\mathrm{rad}\right)^2} = 61 \pm 22
  \, \kms \ .
\end{equation}
This value for the three-dimensional speed of VFTS682 with respect the
surrounding stars make it a ``bona fide'' runaway star.

\subsection{Does it come from the R136 cluster?}
\label{sec:r136_origin}

The inset plot in \Figref{fig:main} shows a zoom in of the region
around R136 which includes VFTS682. To test the hypothesis that the
star was indeed ejected from this region, we check the
orientation of its proper motion relative to the group of stars
selected to represent the R136 cluster in \Secref{sec:sample}. We
further consider also the kinematic age of the star. The error cone, obtained by
propagating the errors on the proper motion components indicates that
R136 is indeed the most likely origin, as expected by \cite{fujii:11,
  banerjee:12}.
